---
title: "Caples_Sat_Veg_Lidar_Comp.Rmd"
author: "DDK"
date: "10/13/2022"
output: html_notebook
---

# Overview of information from satellite, on-the-ground, and Lidar data from the Caples Creek Watershed

```{r setup, echo=FALSE}
knitr::opts_chunk$set(out.extra = '', eval = TRUE, echo = TRUE)
opts <- options(knitr.kable.NA="")

library(cowplot)
library(ggpubr)
library(googledrive)
library(gridExtra)
library(here)
library(kableExtra)
library(leaflet)
library(mapview)
library(patchwork)
library(raster)
library(rgdal)
library(sf)
library(tidyverse)
```

## Introduction

For the Caples Creek Ecological Restoration study we have three types of data:

1.  Remote sensing data and derived products.
2.  On-the-ground vegetation plot measurements.
3.  LIDAR data

To characterize the Caples Creek Watershed before and after both the 2019 Caples fire and the 2021 Caldor fire it is important to do three things:

1.  Ensure the accuracy and precision of the measurement types.
2.  Using this information:
    1.  establish a baseline of data describing relevant variables.

    2.  measure change in these variables across the two fire events while controlling for biological variation not due to fire (growth, mortality) at the same time as minimizing systematic bias.

For basic satellite measurements derived from the [California Forest Observatory](https://forestobservatory.com/), as well as data products from the Rapid Assessment of Vegetation Condition after Wildfire ([RAVG](https://burnseverity.cr.usgs.gov/ravg/)) program.

All data processed in two steps, if starting from scratch, run the following scripts:

```{r}
#source("read_adjust_combine_rasters.R")
#source("read_points_output_data.R")
```
#Map of points

Note both are reliant on 'on-the-ground' locations of sample points for vegetation plot points and avian survey points. So an important step is to ensure those are accurately represented. We have the original UTM 10N (NAD 83) 400m grid provided by Estes (pers. comm). The *avian points* were chosen from those points on this grid and therefore all have Northings of pattern "42###41.3..." and an Easting of "74##46.3..." any *avian points* not on those grid points would be incorrect. *Vegetation* *plot points* were located by first finding one of the local grid points and sometimes moving the vegetation sample point several to less than 30m away to avoid obstacles or better intersect with forest. We will revisit this latter effort later. For the remaining of this analysis, we will utilize the latest confirmed data points from Estes (see [Caples_CSE_Plot_UTM_NAD83_Zone10](https://docs.google.com/spreadsheets/d/174WsPpBfU8IuW2GgsvsiXA-JmNBpG43-6__DmjaMRYA/edit#gid=1008187569)). Inspection of the map for some points suggest that a few vegetation point locations are not correctly located (e.g. Veg #6, Avian #576 which we know to be downslope).  We note that we should fix this issue which will impact only *when we compare* remote sensing data with data collected on the ground. For now, moving along we next visualize the main variables.

```{r 'map of points'}
library(mapview)
mapviewOptions(fgb = FALSE, georaster = FALSE)
#both_points_v from read_points_output_data.R line 198
both_points_v %>% 
  select(veg_point, avian_point, geometry.vegetation) %>% 
  mutate(comb_point = paste0(veg_point,"_", avian_point)) %>%
  mapview(., zcol = "comb_point")
```
##Before After data at the landscape level

### Large Trees
Large trees are defined as the fraction of the area with trees greater than a particular height. In this case we chose 22m cutoff from local dbh vs. height relationships.  Did the large trees change in their study area wide distribution?
#### Large trees at the landscape level
```{r}
#RAVG labels
RAVG_labels <- c("unburned 0", "low (0,1]", "moderate (1,3]", "high (3,4]")

# ggplot faceted histogram large trees versus RAVG data
tall_forest_variables_4ha %>% #from read_points_output_data.R line 1015 
  filter(grepl("LT(.*)Ht|RAVGcbi4", var), Year != 2019, scale == "4ha") %>%
  pivot_wider(
    names_from = sum_fn:Year,
    # names_glue ="{var}_{.value}", #printf('{%s}_{%s}_{%s', sum_fn, scale, Year
    values_from = value
  ) %>%
  pivot_longer(
    cols = contains("Perc"),
    names_to = c("sum_fn", "var", "Year"),
    names_pattern = "(.*)_(.*)_(.*)",
    values_to = "value",
    values_drop_na = TRUE
  ) %>%
  filter(grepl("LTg22mHt", var)) %>%
  mutate(RAVG =cut(mean_RAVGcbi4_20182019, breaks = c(-1, 0, 1, 3, 4), labels = RAVG_labels)) -> tfv_input
#labels = c("0", "(0,1]", "(1,2]", "(2,3]", "(3,4]")
# plot the input
ggplot(tfv_input, aes(value, group = var)) +
  geom_histogram(aes(y = stat(density)), bins = 15) +
  # scale_y_continuous(labels = var) +
  facet_grid(RAVG ~ Year) +
  ggtitle(c("Large Trees (height >= 22m) versus burn severity (RAVG) before and after fire")) +
  xlab("Fraction of 4ha with tree height >= 22m") +
  ylab("Number of plots in each RAVG level")
```
#### Plot level before after large trees

The first thing we noted was that the fraction of 4ha with large trees rose considerably especially in the locations with no fire.  This was problematic. It suggested we look at the repeatability of the tree-heights derived from SALO data.
```{r}
# plot comparing 2018 to 2020
# ltplotinput replicated from "eda_forest_data_points.R" lines 87-97

#plot each year against each other
tall_forest_variables_4ha %>%
  filter(grepl("LT(.*)Ht|RAVGcbi4", var), Year != 2019, scale == "4ha") %>%  # View()
  pivot_wider(
    names_from = sum_fn:Year,
    # names_glue ="{var}_{.value}", #printf('{%s}_{%s}_{%s', sum_fn, scale, Year
    values_from = value
  ) %>% 
  mutate(RAVG =cut(mean_RAVGcbi4_20182019, breaks = c(-1, 0, 1, 3, 4), RAVG_labels)) -> ltplotinput 


ltplotinput$RAVG <- ordered(ltplotinput$RAVG, levels = rev(RAVG_labels))

scale_color_manual_values <- c("high (3,4]" = "black",
                                "moderate (1,3]" = "red",
                                "low (0,1]" = "orange",
                                "unburned 0" = "green")

ggplot(ltplotinput, aes(Perc_LTg22mHt_2018, Perc_LTg22mHt_2020, color = RAVG)) +
  geom_point() +
  xlim(0, .55) +
  ylim(0, .55) +
  coord_equal() +
  theme(aspect.ratio=1) +
  geom_abline(intercept = 0, slope = 1) +
  scale_color_manual(values = scale_color_manual_values) +
  # scale_y_continuous(labels = var) +
  # facet_grid(~RAVG) +
  ggtitle(c("Large Trees (height >= 22m) versus burn severity (RAVG) before and after fire")) +
  xlab("Fraction of 4ha with tree height >= 22m in 2018") +
  ylab("Fraction of 4ha with tree height >= 22m in 2020") 
  
ggplot(ltplotinput, aes(Perc_LTg22mHt_2018, Perc_LTg22mHt_2020, color = RAVG)) +
  geom_point() +
  xlim(0, .55) +
  ylim(0, .55) +
  coord_equal() +
  theme(aspect.ratio=1) +
  geom_abline(intercept = 0, slope = 1) +
  scale_color_manual(values = scale_color_manual_values) +
  # scale_y_continuous(labels = var) +
  facet_grid(~RAVG) +
  ggtitle(c("Large Trees (height >= 22m) versus burn severity (RAVG) before and after fire")) +
  xlab("Fraction of 4ha with tree height >= 22m in 2018") +
  ylab("Fraction of 4ha with tree height >= 22m in 2020") 

```
#### Plot level height and cover

What about mean height and mean cover variables?
```{r}
# wide4havars from read_points_output_data.R
#these delta's must be calculated on a pixel by pixel basis then look at the central tendency [ ]
wide4havars$deltacov <- (wide4havars$mean_CanopyCover_2020_4ha-wide4havars$mean_CanopyCover_2018_4ha)
wide4havars$deltaht <- (wide4havars$mean_CanopyHeight_2020_4ha-wide4havars$mean_CanopyHeight_2018_4ha)
wide4havars$deltabsht <- (wide4havars$mean_CanopyBaseHeight_2020_4ha-wide4havars$mean_CanopyBaseHeight_2018_4ha)
wide4havars$deltabd <- (wide4havars$mean_CanopyBulkDensity_2020_4ha-wide4havars$mean_CanopyBulkDensity_2018_4ha)
wide4havars$deltalc <- (wide4havars$mean_CanopyLayerCount_2020_4ha-wide4havars$mean_CanopyLayerCount_2018_4ha)
wide4havars$deltalf <- (wide4havars$mean_LadderFuelDensity_2020_4ha-wide4havars$mean_LadderFuelDensity_2018_4ha)
wide4havars$deltasf <- (wide4havars$mean_SurfaceFuels_2020_4ha-wide4havars$mean_SurfaceFuels_2018_4ha)

wide4havars %>%
  mutate(RAVG =cut(mean_RAVGcbi4_20182019_4ha, breaks = c(-1, 0, 1, 3, 4), labels = RAVG_labels)) -> wide4havars

wide4havars$RAVG <- ordered(wide4havars$RAVG, levels = rev(RAVG_labels))

wide4havars %>%
  ggplot(aes(x = avian_point, y = deltaht, color = RAVG))+
  geom_point() +
  geom_abline(intercept = 0, slope = 0) +
  scale_color_manual(values = scale_color_manual_values) +
  ggtitle("change in mean canopy height 2020-2018") +
  ylab("change in mean canopy height 4ha") +
  facet_wrap(~RAVG)

wide4havars %>%
  ggplot(aes(x = avian_point, y = deltacov, color = RAVG))+
  geom_point() +
  geom_abline(intercept = 0, slope = 0) +
  scale_color_manual(values = scale_color_manual_values) +
  ggtitle("change in mean canopy cover 2020-2018") +
  ylab("change in mean canopy cover 4ha") +
  facet_wrap(~RAVG)

```

#### Plot level bivariate height & cover
```{r}

wide4havars %>%
  ggplot(aes(x = mean_CanopyHeight_2018_4ha, y = mean_CanopyHeight_2020_4ha, color = RAVG)) +
  geom_point() +
  stat_summary(fun.data=mean_cl_normal) +
  geom_smooth(method='lm', formula= y~x) +
  geom_abline(intercept = 0, slope = 1) +
  coord_fixed() +
  theme(aspect.ratio=1) +
  scale_color_manual(values = scale_color_manual_values) +
  ggtitle(c("Relationship between canopy variables vs \nburn severity (RAVG) before and after fire")) +
  xlab("Mean canopy height 4ha (2018)") +
  ylab("Mean canopy height 4ha (2020)") +
  facet_wrap(~RAVG) 

wide4havars %>%
  ggplot(aes(x = mean_CanopyCover_2018_4ha, y = mean_CanopyCover_2020_4ha, color = RAVG)) +
  geom_point() +
  stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x) +
  geom_abline(intercept = 0, slope = 1) +
  coord_fixed() +
  theme(aspect.ratio=1) +
  scale_color_manual(values = scale_color_manual_values) +
  ggtitle(c("Relationship between canopy variables vs \nburn severity (RAVG) before and after fire")) +
  xlab("Mean canopy cover 4ha (2018)") +
  ylab("Mean canopy cover 4ha (2020)") +
  facet_wrap(~RAVG) 

wide4havars %>%
  ggplot(aes(x = deltaht, y = deltacov, color = RAVG)) +
  geom_point() +
  stat_summary(fun.data=mean_cl_normal) + 
  geom_smooth(method='lm', formula= y~x) +
  geom_abline(intercept = 0, slope = 1) +
  coord_fixed() +
  theme(aspect.ratio=1) +
  scale_color_manual(values = scale_color_manual_values) +
  ggtitle(c("Relationship between canopy variables vs \nburn severity (RAVG) before and after fire")) +
  xlab("Change in height (2020-2018)") +
  ylab("Change in cover (2020-2018)") +
  facet_wrap(~RAVG) 



```
# Large tree analysis discussion: full-model (height, year, plot, slope, intercept)

Using the pixel-level data set for each point Canopy Height to large tree analysis.  For each plot, we show Tree Height 2018 (x) vs 2020 (y), sorted by RAVG, colored by RAVG class, with a separate plot level regression line (slope and intercept) and the LARGE TREE cutoffs shown, to discuss the value of utilizing 'correction factors' for large trees for 2018 vs 2020.

```{r}
#depends on running build_pixel_level_data.R
# source(here("spatial","src", "build_pixel_level_data.R"))
# GGPLOT of Canopy Height analysis
# each plot, sorted by RAVG, colored by RAVG class, 2018 (x) vs 2020 (y)
# Separate slope and intercept
# for model dsicussion
pixel_level_db_ch_cc %>% 
    filter(avian_point != 0) %>%
    # filter(avian_point %in% c(841, 454, 490,576, 1057, 1072)) %>%
    ggplot(aes(x = canopy_height_18, y = canopy_height_20)) +
    # geom_point(pch = ".") +
    # stat_summary(fun.data=mean_cl_normal) + 
    geom_point(aes(size = coverage_fraction, col=RAVG_category), 
               alpha = 0.42, 
               position = jitter, 
               shape = 1, 
               # fill = "tan", 
               stroke = .5) +
    scale_color_manual(values=pt_colors) +
    scale_size(range = c(0,1)) +
      geom_smooth(method="lm",
                formula= y~x,
                mapping = aes(weight = coverage_fraction),
                color = "purple",
                lwd = 0.5,
                show.legend = FALSE) +
    geom_abline(intercept = 0, slope = 1, lty = 2, col = "blue") +
    geom_vline(xintercept = 22, lty = 3, col = "black") +
    geom_hline(yintercept = 22, lty = 3, col = "black") +
    coord_fixed(ratio = 1, xlim = c(0,40), ylim = c(0,40)) +
      # xlim(0, 40) +
      # ylim(0, 40) +
    # theme(aspect.ratio=1) +
    theme(legend.position="none") +
    facet_wrap(~avian_point) + 
    # geom_vline(aes(xintercept = mean, group = avian_point), colour = 'red') +
    #theme(strip.text = element_text(size = rel(3.0), vjust = -4.0), 
    #     panel.spacing.y = unit(-2, "lines")) # +
     theme(strip.text = element_blank(), 
         panel.spacing.y = unit(0.1, "lines"),
         panel.spacing.x = unit(0.1, "lines")) +
         theme(axis.text.x=element_text(size=7, angle = 90)) +
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
         theme(axis.text.y=element_text(size=7)) +
  theme(axis.line = element_line(color='black'),
        # plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank())
```
# Canopy Height Pixel by pixel after - before comparisons - base-R

```{r}
# Pixel by pixel comparisons around particular points 2018 vs 2020
# Depends on canopy_imgStack
# install.packages("naturalsort")
library(naturalsort)
#generate data for CH
exactextractr::exact_extract(canopy_imgStack$CaplesCanopyHeight2018, wldf_100) -> test_ch18 # %>% head()
exactextractr::exact_extract(canopy_imgStack$CaplesCanopyHeight2020, wldf_100) -> test_ch20 # %>% head()

#and CC
exactextractr::exact_extract(canopy_imgStack$CaplesCanopyCover2018, wldf_100) -> test_cc18 # %>% head()
exactextractr::exact_extract(canopy_imgStack$CaplesCanopyCover2020, wldf_100) -> test_cc20 # %>% head()

tibble::as_tibble(test_ch18, .rows = 441, .name_repair = "universal") %>% 
  tibble::rownames_to_column() %>% 
  pivot_longer(cols = -rowname) %>% 
  mutate(pixel_name = rowname) %>% 
  mutate(veg_point = rep(wldf_100$veg_point, 441)) %>% 
  mutate(avian_point = rep(wldf_100$avian_point, 441)) %>%
  mutate(point_idx = name) %>% 
  mutate(canopy_height_18 = value) %>% 
  select(c(7,5:6,4,8)) %>% 
  arrange(point_idx) -> test_ch18_long

tibble::as_tibble(test_ch20, .rows = 441, .name_repair = "universal") %>% 
  tibble::rownames_to_column() %>% 
  pivot_longer(cols = -rowname) %>% 
  mutate(pixel_name = rowname) %>% 
  mutate(veg_point = rep(wldf_100$veg_point, 441)) %>% 
  mutate(avian_point = rep(wldf_100$avian_point, 441)) %>%
  mutate(point_idx = name) %>% 
  mutate(canopy_height_20 = value) %>% 
  select(c(7,5:6,4,8)) %>% 
  arrange(point_idx) -> test_ch20_long

tibble::as_tibble(test_cc18, .rows = 441, .name_repair = "universal") %>% 
  tibble::rownames_to_column() %>% 
  pivot_longer(cols = -rowname) %>% 
  mutate(pixel_name = rowname) %>% 
  mutate(veg_point = rep(wldf_100$veg_point, 441)) %>% 
  mutate(avian_point = rep(wldf_100$avian_point, 441)) %>%
  mutate(point_idx = name) %>% 
  mutate(canopy_cover_18 = value) %>% 
  select(c(7,5:6,4,8)) %>% 
  arrange(point_idx) -> test_cc18_long

tibble::as_tibble(test_cc20, .rows = 441, .name_repair = "universal") %>% 
  tibble::rownames_to_column() %>% 
  pivot_longer(cols = -rowname) %>% 
  mutate(pixel_name = rowname) %>% 
  mutate(veg_point = rep(wldf_100$veg_point, 441)) %>% 
  mutate(avian_point = rep(wldf_100$avian_point, 441)) %>%
  mutate(point_idx = name) %>% 
  mutate(canopy_cover_20 = value) %>% 
  select(c(7,5:6,4,8)) %>% 
  arrange(point_idx) -> test_cc20_long

bind_cols(test_ch18_long,
          test_ch20_long$canopy_height_20,
          test_cc18_long$canopy_cover_18,
          test_cc20_long$canopy_cover_20) %>% 
  unnest(cols = c(canopy_height_18)) %>% 
  select(c(1:4,6,5,7,9,11)) -> pixel_level_db_ch_cc

colnames(pixel_level_db_ch_cc) <- c("point_idx",
                                      "veg_point",
                                      "avian_point",
                                      "pixel_name",
                                      "coverage_fraction",
                                      "canopy_height_18",
                                      "canopy_height_20",
                                      "canopy_cover_18",
                                      "canopy_cover_20")

pixel_level_db_ch_cc %>%
  View()



avian_points_sorted <- naturalsort::naturalsort(wldf_100$avian_point[wldf_100$avian_point != 0])

for(i in avian_points_sorted){
avian_point_number <- i
rowidno <- as.numeric(dimnames(wldf_100)[[1]][wldf_100$avian_point==avian_point_number])

par(mfrow=c(1,1))
par(pty="s")
par(mar=c(4,4,3,2))
plot(jitter(test_ch18[[rowidno]][,1]), jitter(test_ch20[[rowidno]][,1]), xlab = "Canopy Height in 2018", ylab = "Canopy Height in 2020",pch = ".", cex = 3)
abline(0,1)
abline(h = 22, v = 22)
abline(lm(test_ch20[[rowidno]][,1]~test_ch18[[rowidno]][,1]), col="red")
title(main=paste0("Pixel by pixel comparison \n for point ", avian_point_number))  #, " 2018 vs 2020"

#abline(lm(test_ch20[[rowidno]][,1] ~ 1 + offset(1* test_ch18[[rowidno]][,1])), col="red")

test_ch18[[rowidno]][,1] >= 22 -> ch2018g22
sum(ch2018g22 * 1)/length(test_ch18[[rowidno]][,1]) -> pctCHg22_2018
text(x = max(test_ch20[[rowidno]][,1])*.7, y = max(test_ch18[[rowidno]][,1])*.2,labels = paste(as.character(round(pctCHg22_2018,2)*100), "% \nlarge tree"))

test_ch20[[rowidno]][,1] >= 22 -> ch2020g22
sum(ch2020g22 * 1)/length(test_ch20[[rowidno]][,1])  -> pctCHg22_2020
text(x = max(test_ch18[[rowidno]][,1])*.2, y = max(test_ch20[[rowidno]][,1])*.9,labels = paste(as.character(round(pctCHg22_2020,2)*100), " % \nlarge tree"))
}

```
# Canopy Cover Pixel by pixel after - before comparisons - base-R

```{r }
for(i in avian_points_sorted){
avian_point_number <- i
rowidno <- as.numeric(dimnames(wldf_100)[[1]][wldf_100$avian_point==avian_point_number])
par(mfrow=c(1,1))
par(pty="s")
par(mar=c(4,4,3,2))
plot(jitter(test_cc18[[rowidno]][,1]), jitter(test_cc20[[rowidno]][,1]), xlab = "Canopy Cover in 2018", ylab = "Canopy Cover in 2020", pch = ".", cex = 3)
abline(0,1)
# abline(h = 22, v = 22)
abline(lm(test_cc20[[rowidno]][,1]~test_cc18[[rowidno]][,1]), col="red")
title(main=paste0("Pixel by pixel comparison \n for point ", avian_point_number))  #, " 2018 vs 2020"

}
```
# Large tree analysis discussion: full-model (height, year, plot, slope, intercept)

Using the pixel-level data set for each point Canopy Height to large tree analysis.  For each plot, we show Tree Height 2018 (x) vs 2020 (y), sorted by RAVG, colored by RAVG class, with a separate plot level regression line (slope and intercept) and the LARGE TREE cutoffs shown, to discuss the value of utilizing 'correction factors' for large trees for 2018 vs 2020.

```{r}
#depends on running build_pixel_level_data.R
# source(here("spatial","src", "build_pixel_level_data.R"))
# GGPLOT of Canopy Height analysis
# each plot, sorted by RAVG, colored by RAVG class, 2018 (x) vs 2020 (y)
# Separate slope and intercept
# for model dsicussion
pixel_level_db_ch_cc %>% 
    filter(avian_point != 0) %>%
    # filter(avian_point %in% c(841, 454, 490,576, 1057, 1072)) %>%
    ggplot(aes(x = canopy_height_18, y = canopy_height_20)) +
    # geom_point(pch = ".") +
    # stat_summary(fun.data=mean_cl_normal) + 
    geom_point(aes(size = coverage_fraction, col=RAVG_category), 
               alpha = 0.42, 
               position = jitter, 
               shape = 1, 
               # fill = "tan", 
               stroke = .5) +
    scale_color_manual(values=pt_colors) +
    scale_size(range = c(0,1)) +
      geom_smooth(method="lm",
                formula= y~x,
                mapping = aes(weight = coverage_fraction),
                color = "purple",
                lwd = 0.5,
                show.legend = FALSE) +
    geom_abline(intercept = 0, slope = 1, lty = 2, col = "blue") +
    geom_vline(xintercept = 22, lty = 3, col = "black") +
    geom_hline(yintercept = 22, lty = 3, col = "black") +
    coord_fixed(ratio = 1, xlim = c(0,40), ylim = c(0,40)) +
      # xlim(0, 40) +
      # ylim(0, 40) +
    # theme(aspect.ratio=1) +
    theme(legend.position="none") +
    facet_wrap(~avian_point) + 
    # geom_vline(aes(xintercept = mean, group = avian_point), colour = 'red') +
    #theme(strip.text = element_text(size = rel(3.0), vjust = -4.0), 
    #     panel.spacing.y = unit(-2, "lines")) # +
     theme(strip.text = element_blank(), 
         panel.spacing.y = unit(0.1, "lines"),
         panel.spacing.x = unit(0.1, "lines")) +
         theme(axis.text.x=element_text(size=7, angle = 90)) +
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
         theme(axis.text.y=element_text(size=7)) +
  theme(axis.line = element_line(color='black'),
        # plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank())
```

